---
title: "Lab 7 Notes"
author: "Alyssa Winchell"
date: "February 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages

```{r include=FALSE}

library(tidyverse)
library(tmap)
library(sf)
library(spatstat)
library(maptools)
library(sp)
library(raster)
library(gstat)

```

Part 1: Hawaii raster practice

```{r}

hi_par <- raster("PAR_CLIM_M.tif")

hi_sst <- raster("SST_LTM.tif")

hi_chl <- raster("CHL_LTM.tif")

par(mfrow = c(1,3))
plot(hi_par)
plot(hi_sst)
plot(hi_chl)

#can call one of these rasters in console and it will give lots of info about it eg resolution, extent, dimensions, CRS...
#hi_sst@extent in console will also give that specific info

```

Reproject to WGS84:

```{r}

#spacing matters for this code
wgs84 <- "+proj=longlat +datumWGS84 +ellps=WGS84 +no_defs"

hi_sst_84 <- projectRaster(hi_sst, crs = wgs84, method = "bilinear")
hi_sst_84@crs

plot(hi_sst_84)
#to change resolution if higher than needed, can resample using agregate

```

Resample: use raster::aggregate()

```{r}

sst_rs <-aggregate(hi_sst, fact = 10)
plot(sst_rs)
#do only when it makes sense but be careful. can be useful to use when just checking if code is working so you don't have to load a lot

```

Crop a raster:

```{r}

hi_sst_84@extent
plot(hi_sst_84)

#create a reasonable bounding box
bounds <- as(extent(-156.2, -154.5, 18.7, 20.4), 'SpatialPolygons')

crs(bounds) <- crs(hi_sst_84) #make bounds' CRS match hi_sst_84 but still keeps its own extent

sst_crop <- crop(hi_sst_84, bounds) #crop the raster by the bounds
plot(sst_crop)

```

Simple raster math:

Nonsensical variable called "tropicality" that is the sum of PAR +SST + 2*ChlA and we want to map that variable

```{r}

hi_par_84 <- projectRaster(hi_par, crs = wgs84, metod = "bilinear") #bilinear for continuous, nearest neighbor for thematic/categorical raster data using methos = "ngm"
#check CRS in console: hi_par_84@crs

hi_chl_84 <- projectRaster(hi_chl, crs = wgs84, metod = "bilinear")

par(mfrow = c(1,3))
plot(hi_par_84)
plot(hi_sst_84)
plot(hi_chl_84)
#they are not in the same scale on legend

```

```{r}

trop <- hi_par_84 + hi_sst_84 + 2*hi_chl_84 #some things are excluded bc there are extent issues (that's what the warning that comes up means)
plot(trop)

```

Now let's look at something using tmap:

```{r}

#make folder that its in the working directory
islands <- read_sf(dsn = 'islands', layer = "Island_boundaries") %>%
  dplyr::select(Island) %>%
  st_simplify(dTolerance = 10) %>%
  st_transform(crs = 4326) #puts it into wgs84 CRS

plot(islands)

```

Use tmap to view

static viewing instead of interactive: mode plot instead of view

```{r}

tmap_mode("plot")

sst_map <- tm_shape(hi_sst_84) +
  tm_raster(title = "Mean Sea Surface Temperature")+
  tm_layout(bg.color = "navyblue",
            legend.position = c("left", "bottom"),
            legend.text.color = "white",
            legend.text.size = 0.5) +
  tm_shape(islands) +
  tm_fill("darkgreen")

tmap_save(sst_map, "HawaiiExample.png", height = 5) #will save it rather than plotting it. don't include if you only want to plot

```

Conditional rasters and masking:

Let's say we have a sensitive species and we're trying to find habitat it might like. We know they like warm water (average >= 25.6 degress) and solar (PAR) below 54. Very rough example, mostly for the Skills.

```{r}

#check to see if deminstions and extent match for your layers and then make them align perfectly
#make extent the same
extent(hi_sst_84) <- extent(hi_par_84)

cr <- raster(nrow = 822, 
             ncol = 1229, 
             xmn = -160.4365, 
             xmx = -154.5373, 
             ymn = 18.7309, 
             ymx = 22.44634) #matches the one we want to resample to

sst_new <- resample(hi_sst_84, cr) #instead of cr, could also just put in the raster that you want it to match directly

#tells you if things match within rasters
compareRaster(sst_new, hi_par_84) #if error, tells you what is different and what to fix

plot(sst_new)
plot(hi_par_84)

#make cropped version for just hawaii

bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')

crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(par_kauai)
plot(sst_kauai)

compareRaster(sst_kauai, par_kauai)

```

Now we only want to isolate regions where temp >= 25.4, PAR < 54

```{r}

par_hab <- par_kauai
par_hab[par_hab >= 54] <- NA
plot(par_hab)

sst_hab <- sst_kauai
sst_hab[sst_hab < 25.4] <- NA
plot(sst_hab)

#find their intersection through masking. raster:: mask
suit_hab <- mask(sst_hab, par_hab)
plot(suit_hab)

```

##Part 2: point pattern analysis

An analysis of red tree voles in Humboldt County

Read in data:

```{r}

voles <- read_sf(dsn = 'redtreevoledata', layer = "ds033") %>%
  dplyr::select(COUNTY) %>%
  filter(COUNTY == "HUM") %>% 
  st_transform(crs = 4326)

plot(voles)

#get humoldt county outline
humboldt <- read_sf(dsn = 'redtreevoledata', layer = "california_county_shape_file") %>%
  filter(NAME == "Humboldt") %>%
  dplyr::select(NAME)

st_crs(humboldt) <- 4326

plot(humboldt)

#plot both together
tm_shape(humboldt) +
  tm_fill() +
  tm_shape(voles) +
  tm_dots(size = 0.2)

#is there a clustering pattern or are they indep?
ggplot() +
  geom_sf(data = humboldt) +
  geom_sf(data = voles) +
  
ggsave("humvoles.png", 
       units = "in", 
       width = 4, 
       height = 6, 
       dpi = 300)

```

Going exploring: pt patterns in a few different ways:

- quadrat analysis
- distance based (neighbor analysis, G-fxn and K-fxn)

```{r}

voles_sp <- as(voles, 'Spatial')
voles_ppp <- as(voles_sp, 'ppp') #point pattern

humboldt_sp <- as(humboldt, "Spatial")
humboldt_win <- as(humboldt_sp, "owin") #as outer window

voles_pb <- ppp(voles_ppp$x, voles_ppp$y, window = humboldt_win) #creates point pattern: requires points and window
plot(voles_pb) #now contained as one spatial unit (pts and window boundary)

```

Quadrat test:

```{r}

voles_qt <- quadrat.test(voles_pb, nx = 5, ny = 10) #gives a warning if n is too small in some of the quadrants

#testing the nul hyp of spatial evennes (although you will hear it called a test for CSR)
voles_qt
#based on these obs, reject null of evenness -> would conclude that these events do not reflect spatial evenness

plot(voles_pb)
plot(voles_qt, add = TRUE, cex = 0.4) #above is expected # of voles per quadrant (less for partial quadrats) if evenness = TRUE and actual below
#this could be manipulated by changing quadrat size-- be careful! Use !) literature values or 2) logical values

```

Plotting kernal densities for spatial data:

```{r}

point_density <- density(voles_pb, sigma = 0.02)
plot(point_density)
#risky bc bandwidth completely changes impression of data. the way they look is therefore completely arbitrary depending on bw choice. MUST ALSO REPORT BW IF USING

vole_raster <- raster(point_density, crs = wgs84)

tm_shape(vole_raster) +
  tm_raster()

```

Nearest Neighbor:

G-function considers the distance of each obs to its nearest neighbor
K-function considers how close all neighboring obs are to an event (concentric circles)

```{r}

r = seq(0, 0.15, by = 0.005)

gfunction <- envelope(voles_pb, fun =  Gest, r = r, nsim = 20)

plot(gfunction$obs ~ gfunction$r, type = "l", col = "black") #obs as a fxn of r
lines(gfunction$theo ~ gfunction$r, type = "l", col = "red") #plots theoretical csr
#black line (our obs) shows higher prop of events exist w a nearest neighbor at closer distances than at csr in red = our obs are more clustered and prob not spatially random

#drawback: oversimplifies relationship of events bc only considers 1 nearest neighbor so we can look at k function or in this case, the l function which is the standardized version

#k/l fxn asks how close all neighbors are to EVERY event in the spatial window

r2 <- seq(0, 0.5, by = 0.05) #needs greater distance of events bc includes all events instead of 1

lfunction <- envelope(voles_pb, fun = Lest, r = r2, nsim = 20, global = TRUE)

plot(lfunction$obs ~ lfunction$r, type = "l", col = "blue")
lines(lfunction$theo ~ lfunction$r, type = "l", col = "red")
#results are also indicative of clustering compared to csr

```

Digggle-Cressie-Loosmore-Ford test of CSR

```{r}

DCLFTest <- dclf.test(voles_pb, nsim = 30) 
DCLFTest
#give less weight that what you see and the g and l test bc if you have enough n will find diff from CSR no matter what bc of your high power

```

##Part 3: Spatial interpolation by kriging

```{r}

ks_rain <- read_csv("KSRain2.csv")
ks_sf  <-  st_as_sf(ks_rain, coords = c("LON", "LAT"), 
                 crs = 4326) #convert foreign obj to sf obj. give it in long then lat order

plot(ks_sf)

# Get county data
ks_counties <- read_sf(dsn = 'KSCounties', layer = "ks_counties_shapefile")
st_crs(ks_counties) = 4326

plot(ks_counties)

tm_shape(ks_counties) +
  tm_fill()+
  tm_shape(ks_sf) +
  tm_dots("AMT", size = 0.5)

```

Convert to spatial file for the rainfall

```{r}

ks_sp <- as_Spatial(ks_sf)

```

Make spatial grid to interp values over

```{r}

lat <- seq(37, 40, length.out = 200)
lon <- seq(-94.6, -102, length.out = 200)

grid <- expand.grid(lon = lon, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326) #make spatial grid out of the vectors and then gives it spatial info
grid_sp <- as_Spatial(grid_sf) #turns into spatial points data frame

```

make a variogram and find variogram model

```{r}

ks_vgm <- variogram(AMT ~ 1 , ks_sp) #~1 means using ordinary kriging

plot(ks_vgm)

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph")) #give reasonalbe sarting points for iterations to come to convergence
#Sph is spherical model but can use others (see key for example and how to check sum of squares for residuals)

plot(ks_vgm, ks_vgm_fit)

```

Now we can do spatial interp:

```{r}

ks_krige <- krige(AMT ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)


```

Visualize it:

```{r}

ks_krige_df <- as.data.frame(ks_krige)

ks_krige_2 <- ks_krige_df %>%
  rename(lon = coords.x1, lat = coords.x2, predicted = var1.pred, err = var1.var)

rain_predict <- st_as_sf(ks_krige_2, coords = c("lon", "lat"), crs = 4326)

# Get Kansas outline to crop: 
ks <- read_sf(dsn = "states", layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(rain_predict)

#crop grid to outline of KS
rain_crop <- st_intersection(rain_predict, ks)
plot(rain_crop)

```

Look at key for plotting kriging outcomes with tmap
